Type-Safe Modular Parsing

### 2017-01-10 paper/sections/PackratParsing.tex
Revised S2

After left-recursion elimination, the structure of grammar is changed, as well as its corresponding parser. In a modular setting, it is possible but unnecessorily compilcated to analyse the grammar and rewrite it when doing extension. Anticipating that every non-terminal has left-recursive rules, even though some do not, is helpful for extension. However, it is inconvenient and introduces extra complexity for representation of grammar and implementation of parser.

@@

\subsection{Packrat Parsing}\label{subsec:packratparsing}
Fortunately some more advanced parsing techniques such as Packrat parsing~\cite{Ford2002} have been developed to address limitations of simple parser combinators. Packrat parsers use memoization to record the result of applying each parser at each position of the input, so that repeated computation is eliminated. Moreover, Packrat parsing supports both direct and indirect left-recursion~\cite{warth2008} theoretically. All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique in our modular parsing library.
\begin{comment}
It is worth mentioning that the choice of parser combinators will not
affect the other parts of our library. One can choose other parser
combinators like Parsec, in cases that the performance and supporting
of left-recursion are not major concerns. A different library can even build a new
\name with fancy features or higher efficiency.
\end{comment}
Scala has a standard parser combinator library\footnote{https://github.com/scala/scala-parser-combinators}
\cite{moors2008parser} which implements Packrat parsers.
The library provides a number of parser combinators, including the longeset match alternative combinator.
Below we present an example to illustrate Scala parsers.
%
%\bruno{So, we make a big fuss identifying the problems with other
%parser combinators in the previous subsection and now we present
%an example that does not suffer from any of the previous problems!
%What makes sense here is an example that uses: a \emph{left-recursive grammar}; the \emph{longest match combinator}; and where we can argue that
%there's no need for the user to do manual backtracking. Just use
%the example with the left-recursive grammar for parsing integers and additions and argue that none of the limitations
%discussed previously applies!
%}

@@

It is straightforward to model ASTs by inheritance and write corresponding parsers for all cases.
Note that a parser has type \lstinline{PackratParser[E]} for some
\lstinline{E}, which indicates the type of results it produces. Figure~\ref{fig:packrat-arith} shows the abstract syntax and
the parser code that generates expressions as ASTs. Meanwhile, pretty-printing shows up as an operation on the AST.

@@

It is worth mentioning that such a grammar is left-recursive in the addition case. Since left-recursion is well supported by Packrat parsers, we do not need extra code for it. Our parser also employs longest match composition by using the combinator \lstinline{|||} in the library. Furthermore, it does not suffer from the backtracking problem, as the memoization technique of Packrat parsing guarantees the efficiency.

>>>>

After left-recursion elimination, the structure of grammar is changed,
as well as its corresponding parser. In a modular setting, it is
possible but unnecessarily compilcated to analyse the grammar and
rewrite it when doing extension. Anticipating that every non-terminal
has left-recursive rules is helpful for extensibility but overkill,
since it is inconvenient and introduces extra complexity for
representation of grammar and implementation of parser.

@@

\subsection{Packrat Parsing}\label{subsec:packratparsing}
Fortunately some more advanced parsing techniques such as Packrat
parsing~\cite{Ford2002} have been developed to address limitations of
simple parser combinators. Packrat parsers use memoization to record
the result of applying each parser at each position of the input, so
that repeated computation is eliminated. Moreover, Packrat parsing, in
theory, supports both direct and indirect left-recursion~\cite{warth2008}. All of these properties are very suitable for modularity, thus we decided to use Packrat parsers as the underlying parsing technique in for modular parsing.
\begin{comment}
It is worth mentioning that the choice of parser combinators will not
affect the other parts of our library. One can choose other parser
combinators like Parsec, in cases that the performance and supporting
of left-recursion are not major concerns. A different library can even build a new
\name with fancy features or higher efficiency.
\end{comment}
Scala has a standard parser combinator library\footnote{https://github.com/scala/scala-parser-combinators}
\cite{moors2008parser} which implements Packrat parsers.
The library provides a number of parser combinators, including the longeset match alternative combinator.
Below we present an example to illustrate Scala parsers.
%
%\bruno{So, we make a big fuss identifying the problems with other
%parser combinators in the previous subsection and now we present
%an example that does not suffer from any of the previous problems!
%What makes sense here is an example that uses: a \emph{left-recursive grammar}; the \emph{longest match combinator}; and where we can argue that
%there's no need for the user to do manual backtracking. Just use
%the example with the left-recursive grammar for parsing integers and additions and argue that none of the limitations
%discussed previously applies!
%}

@@

It is straightforward to model ASTs by inheritance and write corresponding parsers for all cases.
Note that a parser has type \lstinline{PackratParser[E]} for some
\lstinline{E}, which indicates the type of results it produces. Figure~\ref{fig:packrat-arith} shows the abstract syntax and
the parser code that generates expressions as ASTs. The AST supports
pretty-printing via the \lstinline{print} method.

@@

It is worth mentioning that the grammar is left-recursive in the addition case. Since left-recursion is well supported by Packrat parsers, we do not need extra code for it. Our parser also employs longest match composition by using the combinator \lstinline{|||} in the library. Furthermore, it does not suffer from the backtracking problem, as the memoization technique of Packrat parsing guarantees the efficiency.
